Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-03-22 doc/paper/paper.Rnw
changes from meeting today

There was substantial heterogeneity among the parameter estimates for the five
published HIV datasets we analysed. Two of the
datasets~\autocite{niculescu2015recent, wang2015targeting} had estimated
$\alpha$ values near unity (\gls{MAP} estimate [95\% \gls{HPD}] 
  \Sexpr{map["niculescu2015", alpha]} 
  [\Sexpr{low["niculescu2015", alpha]}-\Sexpr{up["niculescu2015", alpha]}]
and
  \Sexpr{map["wang2015", alpha]} 
  [\Sexpr{low["wang2015", alpha]}-\Sexpr{up["wang2015", alpha]}]). Another two
datasets~\autocite{li2015hiv, cuevas2009hiv} had lower estimated values and wider
\gls{HPD} intervals
  (\Sexpr{map["li2015", alpha]} 
  [\Sexpr{low["li2015", alpha]}-\Sexpr{up["li2015", alpha]}]
and
  \Sexpr{map["cuevas2009", alpha]} 
  [\Sexpr{low["cuevas2009", alpha]}-\Sexpr{up["cuevas2009", alpha]}]). 
The \textcite{novitsky2014impact} data had an extremely low estimated $\alpha$
and a very wide \gls{HPD} interval
  \Sexpr{map["novitsky2014", alpha]} 
  [\Sexpr{low["novitsky2014", alpha]}-\Sexpr{up["novitsky2014", alpha]}]). 
For all the datasets except \textcite{novitsky2014impact}, estimated values of
$I$ were below 2000, with narrow \gls{HPD} intervals around two of the
datasets
  (\textcite{cuevas2009hiv}, \Sexpr{map["cuevas2009", I]} 
  [\Sexpr{low["cuevas2009", I]}-\Sexpr{up["cuevas2009", I]}];
   \textcite{niculescu2015recent}, \Sexpr{map["niculescu2015", I]}
  [\Sexpr{low["niculescu2015", I]}-\Sexpr{up["niculescu2015", I]}])
and wider intervals around the other two
  (\textcite{li2015hiv}, \Sexpr{map["li2015", I]} 
  [\Sexpr{low["li2015", I]}-\Sexpr{up["li2015", I]}];
   \textcite{wang2015targeting}, \Sexpr{map["wang2015", I]}
  [\Sexpr{low["wang2015", I]}-\Sexpr{up["wang2015", I]}]).
The \textcite{novitsky2014impact} data was again the outlier, with a very high
estimated $I$, and \gls{HPD} interval spanning almost the entire prior region
  (\Sexpr{map["novitsky2014", I]} 
  [\Sexpr{low["novitsky2014", I]}-\Sexpr{up["novitsky2014", I]}]).
Little information was gleaned about the $m$ parameter, with the \gls{HPD}
interval occuping the entire prior region for all datasets. The estimates of
$N$ were similarly uninformative, with the exception that the point estimate
for the \textcite{wang2015targeting} data was smaller
  (\Sexpr{map["wang2015", N]})
than the estimates for other datasets.

\begin{figure}[ht]
  \centering
  \includegraphics{realdata-hpd.pdf}
  \caption{
    Point estimates and 95\% \gls{HPD} intervals for \gls{BA} network
    parameters of published HIV datasets, obtained with kernel-ABC.
  }
\end{figure}

@@

Contact networks can have a strong influence on epidemic progression, and are
potentially useful as a public health tool. Despite this, few methods exist for
investigating contact network parameters in a phylodynamic framework.
Kernel-ABC is a model-agnostic method which can be used to investigate any
quantity that affects tree shape. In this work, we developed a kernel-ABC-based
method to infer the parameters of a contact network model. The method is
general, and could be applied to any model from which contact networks can be
simulated. We demonstrated the method on the \gls{BA} model, which is a simple 
model incorprorating the preferential attachment feature commonly seen in
real-world networks.

By training a kernel-SVM classifier, we found that the $\alpha$ and $I$
parameters, representing preferential attachment power and number of infected
nodes, had a strong influence on tree shape. This was reflected in the relative
accuracy of the kernel-ABC estimates of these parameters. On the other hand,
the $m$ and $N$ parameters did not produce much variation in tree shape,
resulting in both poorly performing classifiers and uninformative kernel-ABC
estimates. The total number of nodes $N$ was almost always significantly
over-estimated. Since the prior on $N$ and $I$ is jointly uniform on a
triangular region ($I \leq N$), there is more prior mass on high $N$ values. In
retrospect, it is unreasonable to expect good estimation of $N$, because adding
more nodes to a \gls{BA} network does not change the edge density or overall
shape. This can be illustrated by imaging that we add a small number of nodes
to a network after the epidemic simulation has already been completed. It is
possible that none of these new nodes attains a connection to any infected
node. Thus, running the simulation again on the new, larger network could
produce the exact same transmission tree as before.

@@

Our investigation of published HIV datasets indicated heterogeneity in the
contact network structures underlying several distinct local epidemics.
\textcite{cuevas2009hiv} studied a group of newly diagnosed individuals in the
Basque Country, Spain. Although the individuals were of mixed risk group, and
therefore unlikely to comprise a single contact network, a high proportion of
them (47\%) grouped into local transmission clusters, allowing for the
possibility of weak preferential attachment dynamics. We estimated a relatively
low attachment power for these data, which is consistent with the sampled
sequences comprising many distinct sub-networks. \textcite{li2015hiv} sampled a
large number of acutely infected \gls{MSM} in Shanghai, China. By using a
patristic distance cutoff~\autocite{poon2014impact}, we identified and analysed
a large cluster. Surprisingly, the estimated attachment power was fairly low
with a large credible interval.

\textcite{niculescu2015recent} studied a recent outbreak among Romainian
\gls{IDU}. The estimated attachment power for this dataset was the highest
among all we considered, at slightly above one. The dataset constituted a
fairly high sampling prevalence of 33\% of newly diagnosed \gls{IDU} infections
over the study period. \textcite{wang2015targeting} sampled acutely infected
\gls{MSM} in Beijing, China, and discovered a high degree of clustering which
indicated that the local \gls{MSM} subnetwork had been deeply sampled. In both
these datasets, the estimated attachment power was close to one.

\textcite{novitsky2013phylogenetic} sampled approximately 44\% of the
HIV-infected individuals in the northern area of Mochudi, Botswana. Additional
sampling in a later study~\autocite{novitsky2014impact} brought the genotyping
coverage up to 70\%. Even with such a high sampling coverage, we did not detect
any large clusters using patristic distance, and therefore chose to analyze a
subtree instead. The estimated attachment power was extremely low, with a very
wide credible interval. The estimated number of infected nodes was also
extremely high, much higher than the HIV-infected population of the town, with
a similarly wide credible interval. Several factors may have contributed to
this result. First, the authors note that the their sample was 75\% female. In
a primarily heterosexual risk environmet, removal of a disproportionate nmuber
of males from the network could obfuscate the true network structure. Second,
the town in question was in close proximity to the country's capital, and the
authors indicated that a high amount of migration takes place between the two
locations. This suggests a reason for the high estimate of $I$: the contact
network includes a much larger group based in the capital city.

When interpreting these results, we caution that the \gls{BA} model is quite
simple and most likely misspecified for these data. In particular, the average
degree of a node in the network is equal to $2m$, and therefore is constrained
to be a multiple of 2. Furthermore, we considered the case $m = 1$, where the
network has no cycles, to be implausable and assigned it zero prior
probability. This forces the average degree to be at least four, which may be
unrealistically high for sexual networks. Additional modelling assumptions,
include the network being connected and static, the epidemic following simple
\gls{SI} dynamics, and the underlying behaviour of all nodes is the same. This
last is particularly problematic, as we showed by simulating a network where
some nodes exhibited a higher attachment power than others. The estimated
attachment power was simply the average of the two values, indicating that,
although we could characterize the network in aggregate, the estimated
parameters could not be said to apply to any individual node. However, we note
that, given the model-agnostic nature of our method, it is possible to fit a
model with heterogeneous node behaviour, and indeed this may prove a fruitful
avenue for future investigations.

Our method itself has of caveats, perhaps the most significant being that it
takes a transmission tree as input. In reality, true transmission trees are not
available and must be approximated, often by way of a viral phylogeny. Although
this has been demonstrated to be a fair
approximation~\autocite[e.g.][]{leitner1996accurate}, and is frequently used in
practice~\autocite[e.g.][]{stadler2013uncovering}, the topologies of a viral
phylogeny and transmission tree can differ
significantly~\autocite{ypma2013relating}. In addition, it is computationally
intensive, taking about a day when run on 20 cores in parallel with the setings
we described in the methods. However, there are no other methods available to
perform the types of estimation we do here, and we therefore believe our method
will be useful to epidemiological researchers interested in the network
structure underlying disease outbreaks.

@@

\begin{figure}[ht]
  \centering
  %\includegraphics{smc_test}
  \caption{Approximation of mixture of Gaussians used by
    \textcite{del2012adaptive} and \textcite{sisson2007sequential} to test
    \gls{SMC}. Solid black line indicates true distribution. Grey shaded area
    shows SMC approximation obtained with our implementation.}
  \label{fig:smctest}
\end{figure}

>>>>

There was substantial heterogeneity among the parameter estimates for the five
published HIV datasets we analysed. Two of the
datasets~\autocite{niculescu2015recent, wang2015targeting} had estimated
$\alpha$ values near unity (\gls{MAP} estimate [95\% \gls{HPD}] 
  \Sexpr{map["niculescu2015", alpha]} 
  [\Sexpr{low["niculescu2015", alpha]}-\Sexpr{up["niculescu2015", alpha]}]
and
  \Sexpr{map["wang2015", alpha]} 
  [\Sexpr{low["wang2015", alpha]}-\Sexpr{up["wang2015", alpha]}]). Another two
datasets~\autocite{li2015hiv, cuevas2009hiv} had lower estimated values and wider
\gls{HPD} intervals
  (\Sexpr{map["li2015", alpha]} 
  [\Sexpr{low["li2015", alpha]}-\Sexpr{up["li2015", alpha]}]
and
  \Sexpr{map["cuevas2009", alpha]} 
  [\Sexpr{low["cuevas2009", alpha]}-\Sexpr{up["cuevas2009", alpha]}]). 
The \textcite{novitsky2014impact} data had an extremely low estimated $\alpha$
and a very wide \gls{HPD} interval
  \Sexpr{map["novitsky2014", alpha]} 
  [\Sexpr{low["novitsky2014", alpha]}-\Sexpr{up["novitsky2014", alpha]}]). 
For all the datasets except \textcite{novitsky2014impact}, estimated values of
$I$ were below 2000, with narrow \gls{HPD} intervals around two of the
datasets
  (\textcite{cuevas2009hiv}, \Sexpr{map["cuevas2009", I]} 
  [\Sexpr{low["cuevas2009", I]}-\Sexpr{up["cuevas2009", I]}];
   \textcite{niculescu2015recent}, \Sexpr{map["niculescu2015", I]}
  [\Sexpr{low["niculescu2015", I]}-\Sexpr{up["niculescu2015", I]}])
and wider intervals around the other two
  (\textcite{li2015hiv}, \Sexpr{map["li2015", I]} 
  [\Sexpr{low["li2015", I]}-\Sexpr{up["li2015", I]}];
   \textcite{wang2015targeting}, \Sexpr{map["wang2015", I]}
  [\Sexpr{low["wang2015", I]}-\Sexpr{up["wang2015", I]}]).
The \textcite{novitsky2014impact} data was again the outlier, with a very high
estimated $I$, and \gls{HPD} interval spanning almost the entire prior region
  (\Sexpr{map["novitsky2014", I]} 
  [\Sexpr{low["novitsky2014", I]}-\Sexpr{up["novitsky2014", I]}]).
Little information was gleaned about the $m$ parameter, with the \gls{HPD}
interval occupying the entire prior region for all datasets. The estimates of
$N$ were similarly uninformative, with the exception that the point estimate
for the \textcite{wang2015targeting} data was smaller
  (\Sexpr{map["wang2015", N]})
than the estimates for other datasets.

\begin{figure}[ht]
  \centering
  \includegraphics{realdata-hpd.pdf}
  \caption{
    Point estimates and 95\% \gls{HPD} intervals for \gls{BA} network
    parameters of published HIV datasets, obtained with kernel-ABC.
  }
  \label{fig:abchpd}
\end{figure}

@@

Contact networks can have a strong influence on epidemic progression, and are
potentially useful as a public health tool~\autocite{wang2015targeting,
little2014using}. Despite this, few methods exist for investigating contact
network parameters in a phylodynamic framework~\autocite{groendyke2011bayesian,
brown2011transmission}. Kernel-ABC is a model-agnostic method which can be used
to investigate any quantity that affects tree shape. In this work, we developed
a kernel-ABC-based method to infer the parameters of a contact network model.
The method is general, and could be applied to any model from which contact
networks can be simulated. We demonstrated the method on the \gls{BA} model,
which is a simple model incorporating the preferential attachment feature
commonly seen in real-world networks.

By training a kernel-SVM classifier, we found that the $\alpha$ and $I$
parameters, representing preferential attachment power and number of infected
nodes, had a strong influence on tree shape. This was reflected in the relative
accuracy of the kernel-ABC estimates of these parameters. The total number of
nodes $N$ had a weak influence on tree shape, which was most prominent when the
epidemic size $I$ and number of sampled tips were both large. The $m$
parameter, representing the number of edges created in the network per vertex,
did not produce much variation in tree shape, resulting in in both poorly
performing classifiers and uninformative kernel-ABC estimates.

$N$ was almost always significantly over-estimated using kernel-ABC. Since the
prior on $N$ and $I$ is jointly uniform on a triangular region ($I \leq N$),
there is more prior mass on high $N$ values. In retrospect, it is unreasonable
to expect good estimation of $N$, because adding more nodes to a \gls{BA}
network does not change the edge density or overall shape. This can be
illustrated by imagining that we add a small number of nodes to a network after
the epidemic simulation has already been completed. It is possible that none of
these new nodes attains a connection to any infected node. Thus, running the
simulation again on the new, larger network could produce the exact same
transmission tree as before.

@@

Our investigation of published HIV datasets indicated heterogeneity in the
contact network structures underlying several distinct local epidemics. The
five datasets analysed fell into three categories (\cref{fig:abchpd}). First,
we estimated a preferential attachment power between 0.5 and 1 for the
epidemics studied by \textcite{cuevas2009hiv} and \textcite{li2015hiv}, with
credible intervals occupying nearly the entire region from 0 to 1.
\citeauthor{cuevas2009hiv} studied a group of newly diagnosed individuals in
the Basque Country, Spain. Although the individuals were of mixed risk groups,
and therefore unlikely to comprise a single contact network, a high proportion
of them (47\%) grouped into local clusters based on genetic distance. The low
estimated attachment power for these data is consistent with the sampled
sequences comprising many distinct sub-networks rather than a single connected
network. \citeauthor{li2015hiv} sampled a large number of acutely infected
\gls{MSM} in Shanghai, China, in which we identified a large cluster from the
phylogeny using a patristic distance cutoff~\autocite{poon2014impact}. The low
attachment power estimated for this dataset was surprising given the high
phylogenetic relatedness of the sequences. It is possible that the number and
diversity of circulating recombinant forms introduced errors into the estimated
viral phylogeny.

For the outbreaks studied by \textcite{niculescu2015recent} and
\textcite{wang2015targeting}, the estimated $\alpha$ was close to one, with a
narrower credible interval than for the other studies.
\citeauthor{niculescu2015recent} studied a recent outbreak among Romanian
\gls{IDU}, while \citeauthor{wang2015targeting} sampled acutely infected
\gls{MSM} in Beijing, China. Both studies discovered a high degree of
phylogenetic relatedness owing to the recent infection times and homogeneous
risk groups of the studied populations. The estimated number of infections for
these datasets were also quite low, although the \gls{HPD} interval for
\citeauthor{wang2015targeting} was much wider than that for
\citeauthor{niculescu2015recent}.

The final studied dataset was an outlier in terms of estimated parameters.
\textcite{novitsky2013phylogenetic} sampled approximately 44\% of the
HIV-infected individuals in the northern area of Mochudi, Botswana. Additional
sampling in a later study~\autocite{novitsky2014impact} brought the genotyping
coverage up to 70\%. Even with such a high sampling coverage, we did not detect
any large clusters using patristic distance, and therefore chose to analyze a
subtree instead. Estimates of $\alpha$ and $N$ both had very wide \gls{HPD}
intervals and were markedly different from the other datasets. The estimated
number of infected nodes was also extremely high, much higher than the
estimated HIV prevalence of the town. Several factors may have contributed to
these results. First, the authors note that the their sample was 75\% female.
In a primarily heterosexual risk environment, removal of a disproportionate
number of males from the network could obfuscate the true network structure.
Second, the town in question was in close proximity to the country's capital,
and the authors indicated that a high amount of migration takes place between
the two locations. This suggests that the contact network may include a much
larger group based in the capital city, which would explain the high estimate
of $I$.

When interpreting these results, we caution that the \gls{BA} model is quite
simple and most likely misspecified for these data. In particular, the average
degree of a node in the network is equal to $2m$, and therefore is constrained
to be a multiple of 2. Furthermore, we considered the case $m = 1$, where the
network has no cycles, to be implausible and assigned it zero prior
probability. This forces the average degree to be at least four, which may be
unrealistically high for sexual networks. Additional modelling assumptions
include the network being connected and static, all transmission rates being
equal, no removal after infection, and identical behaviour of all nodes. This
last is particularly problematic, as we showed by simulating a network where
some nodes exhibited a higher attachment power than others. The estimated
attachment power was simply the average of the two values, indicating that,
although we could characterize the network in aggregate, the estimated
parameters could not be said to apply to any individual node. Despite these
issues, we felt it was best to demonstrate the method first on a simple model.
It is possible to fit more complex models which address some of these issues,
such as one incorporating heterogeneous node behaviour. This may prove a
fruitful avenue for future investigations.

Our method has a number of caveats, perhaps the most significant being that it
takes a transmission tree as input. In reality, true transmission trees are not
available and must be approximated, often by way of a viral phylogeny. Although
this has been demonstrated to be a fair
approximation~\autocite[e.g.][]{leitner1996accurate}, and is frequently used in
practice~\autocite[e.g.][]{stadler2013uncovering}, the topologies of a viral
phylogeny and transmission tree can differ
significantly~\autocite{ypma2013relating}. In addition, the \gls{ABC}-\gls{SMC}
algorithm is computationally intensive, taking about a day when run on 20 cores
in parallel with the settings we described in the methods. Nevertheless, our
method is potentially useful to epidemiological researchers interested in the
general characteristics of the network structure underlying disease outbreaks.

@@

\begin{figure}[ht]
  \centering
  \includegraphics{smc-test.pdf}
  \caption{Approximation of mixture of Gaussians used by
    \textcite{del2012adaptive} and \textcite{sisson2007sequential} to test
    \gls{SMC}. Solid black line indicates true distribution. Grey shaded area
    shows SMC approximation obtained with our implementation.}
  \label{fig:smctest}
\end{figure}

### 2016-03-22 doc/paper/paper.tex
changes from meeting today

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The \gls{MAP} [95\%
HPD] estimates for each parameter were: 
$\alpha$, 
  1 [0.07-
                           1.07];
$I$,
  5.37 [2.05-
                       5.82];
$m$,
  \ensuremath{1.01\times 10^{4}} [\ensuremath{3.06\times 10^{3}}-
                              \ensuremath{1.5\times 10^{4}}];
$N$,
  \ensuremath{1.18\times 10^{3}} [505.71-
                       \ensuremath{1.62\times 10^{3}}].

To test the effect of sampling bias, we sampled one transmission tree in a
peer-driven fashion, where the probability to sample a node was twice as high
if one of it's peers had already been sampled. The parameters for this
experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and $I$ = 2000. The
estimated values were
$\alpha$, 
  0.46 [0.03-
                           0.83];
$I$,
  3.53 [2.18-
                       5.9];
$m$,
  \ensuremath{1.07\times 10^{4}} [\ensuremath{3.11\times 10^{3}}-
                              \ensuremath{1.47\times 10^{4}}];
$N$,
  \ensuremath{2.46\times 10^{3}} [\ensuremath{1.28\times 10^{3}}-
                       \ensuremath{3.88\times 10^{3}}].

@@

There was substantial heterogeneity among the parameter estimates for the five
published HIV datasets we analysed. Two of the
datasets~\autocite{niculescu2015recent, wang2015targeting} had estimated
$\alpha$ values near unity (\gls{MAP} estimate [95\% \gls{HPD}] 
  1.06 
  [0.63-1.27]
and
  1 
  [0.41-1.16]). Another two
datasets~\autocite{li2015hiv, cuevas2009hiv} had lower estimated values and wider
\gls{HPD} intervals
  (0.77 
  [0.01-1.03]
and
  0.66 
  [0.03-0.84]). 
The \textcite{novitsky2014impact} data had an extremely low estimated $\alpha$
and a very wide \gls{HPD} interval
  0.17 
  [0.04-1.39]). 
For all the datasets except \textcite{novitsky2014impact}, estimated values of
$I$ were below 2000, with narrow \gls{HPD} intervals around two of the
datasets
  (\textcite{cuevas2009hiv}, 880.52 
  [290.7-\ensuremath{1.51\times 10^{3}}];
   \textcite{niculescu2015recent}, 175.05
  [138.74-454.51])
and wider intervals around the other two
  (\textcite{li2015hiv}, \ensuremath{1.59\times 10^{3}} 
  [284.01-\ensuremath{3.81\times 10^{3}}];
   \textcite{wang2015targeting}, 651.72
  [268.94-\ensuremath{4.24\times 10^{3}}]).
The \textcite{novitsky2014impact} data was again the outlier, with a very high
estimated $I$, and \gls{HPD} interval spanning almost the entire prior region
  (\ensuremath{7.55\times 10^{3}} 
  [228.96-\ensuremath{8.92\times 10^{3}}]).
Little information was gleaned about the $m$ parameter, with the \gls{HPD}
interval occuping the entire prior region for all datasets. The estimates of
$N$ were similarly uninformative, with the exception that the point estimate
for the \textcite{wang2015targeting} data was smaller
  (\ensuremath{5.84\times 10^{3}})
than the estimates for other datasets.

\begin{figure}[ht]
  \centering
  \includegraphics{realdata-hpd.pdf}
  \caption{
    Point estimates and 95\% \gls{HPD} intervals for \gls{BA} network
    parameters of published HIV datasets, obtained with kernel-ABC.
  }
\end{figure}

>>>>

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The \gls{MAP} [95\%
HPD] estimates for each parameter were: 
$\alpha$, 
  1 [0.75-
                           1.12];
$I$,
  5.24 [2.05-
                       5.81];
$m$,
  \ensuremath{1.04\times 10^{4}} [\ensuremath{2.27\times 10^{3}}-
                              \ensuremath{1.5\times 10^{4}}];
$N$,
  \ensuremath{1.17\times 10^{3}} [502.34-
                       \ensuremath{1.58\times 10^{3}}].

To test the effect of sampling bias, we sampled one transmission tree in a
peer-driven fashion, where the probability to sample a node was twice as high
if one of it's peers had already been sampled. The parameters for this
experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and $I$ = 2000. The
estimated values were
$\alpha$, 
  0.46 [0.03-
                           0.83];
$I$,
  3.47 [2.18-
                       5.9];
$m$,
  \ensuremath{1.08\times 10^{4}} [\ensuremath{3.11\times 10^{3}}-
                              \ensuremath{1.48\times 10^{4}}];
$N$,
  \ensuremath{2.94\times 10^{3}} [\ensuremath{1.28\times 10^{3}}-
                       \ensuremath{3.94\times 10^{3}}].

@@

There was substantial heterogeneity among the parameter estimates for the five
published HIV datasets we analysed. Two of the
datasets~\autocite{niculescu2015recent, wang2015targeting} had estimated
$\alpha$ values near unity (\gls{MAP} estimate [95\% \gls{HPD}] 
  1.06 
  [0.63-1.27]
and
  1 
  [0.41-1.16]). Another two
datasets~\autocite{li2015hiv, cuevas2009hiv} had lower estimated values and wider
\gls{HPD} intervals
  (0.77 
  [0.01-1.03]
and
  0.66 
  [0.03-0.84]). 
The \textcite{novitsky2014impact} data had an extremely low estimated $\alpha$
and a very wide \gls{HPD} interval
  0.17 
  [0.04-1.39]). 
For all the datasets except \textcite{novitsky2014impact}, estimated values of
$I$ were below 2000, with narrow \gls{HPD} intervals around two of the
datasets
  (\textcite{cuevas2009hiv}, 880.52 
  [290.7-\ensuremath{1.51\times 10^{3}}];
   \textcite{niculescu2015recent}, 175.05
  [138.74-454.51])
and wider intervals around the other two
  (\textcite{li2015hiv}, \ensuremath{1.59\times 10^{3}} 
  [284.01-\ensuremath{3.81\times 10^{3}}];
   \textcite{wang2015targeting}, 651.72
  [268.94-\ensuremath{4.24\times 10^{3}}]).
The \textcite{novitsky2014impact} data was again the outlier, with a very high
estimated $I$, and \gls{HPD} interval spanning almost the entire prior region
  (\ensuremath{7.55\times 10^{3}} 
  [228.96-\ensuremath{8.92\times 10^{3}}]).
Little information was gleaned about the $m$ parameter, with the \gls{HPD}
interval occupying the entire prior region for all datasets. The estimates of
$N$ were similarly uninformative, with the exception that the point estimate
for the \textcite{wang2015targeting} data was smaller
  (\ensuremath{5.84\times 10^{3}})
than the estimates for other datasets.

\begin{figure}[ht]
  \centering
  \includegraphics{realdata-hpd.pdf}
  \caption{
    Point estimates and 95\% \gls{HPD} intervals for \gls{BA} network
    parameters of published HIV datasets, obtained with kernel-ABC.
  }
  \label{fig:abchpd}
\end{figure}
